---
title: "p8105_hw1_gw2383"
output:
  html_document:
    toc: true
    toc_float: true
author: Guojing Wu
data: 2018-09-20
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

***

## Problem 1

### step 1

We start with creating a data frame:

```{r}
prob1_df = tibble (
  vec_runif = runif(10, min = 0, max = 5), 
  vec_logical = c(vec_runif > 2), # judge the vec_runif and reture True or False
  vec_char = c("p8105", "data", "science", "is", "great", 
               "my", "name", "is", "guojing", "wu"), 
  vec_factor = factor(c(1, 2, 3, 1, 2, 3, 1, 2, 3, 1))
)

# show the result
prob1_df
```

### step 2

Calculate the mean of each variable

<br/>

```{r}
mean(prob1_df$vec_runif)
```  
* Since vec_runif is numerical, we can calculate the mean value.

<br/>

```{r}
mean(prob1_df$vec_logical)
```
* R regard the logical value true as 1 and false as 0, so we can also calculate the mean value.

<br/>

```{r}
mean(prob1_df$vec_char)
```
* The characters are not numerical, so we can't calculate mean value.

<br/>

```{r}
mean(prob1_df$vec_factor)
```
* The characters are not numerical, so we can't calculate mean value.

<br/>
