---
title: "p8105_hw1_gw2383"
output:
  html_document:
    toc: true
    toc_float: true
author: Guojing Wu
data: 2018-09-20
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

***

## Problem 1

### step 1

**We start with creating a data frame:**

```{r}
prob1_df = tibble (
  vec_runif = runif(10, min = 0, max = 5), 
  vec_logical = c(vec_runif > 2), # judge the vec_runif and reture True or False
  vec_char = c("p8105", "data", "science", "is", "great", 
               "my", "name", "is", "guojing", "wu"), 
  vec_factor = factor(c(1, 2, 3, 1, 2, 3, 1, 2, 3, 1))
)

# show the result
prob1_df
```

### step 2

**Calculate the mean of each variable**

```{r}
mean(prob1_df$vec_runif)
```  
* Since vec_runif is numerical, we can calculate the mean value.

<br/>

```{r}
mean(prob1_df$vec_logical)
```
* R regard the logical value true as 1 and false as 0, so we can also calculate the mean value.

<br/>

```{r}
mean(prob1_df$vec_char)
```
* The characters are not numerical, so we can't calculate mean value.

<br/>

```{r}
mean(prob1_df$vec_factor)
```
* The characters are not numerical, so we can't calculate mean value.

### step 3

**Applies the `as.numeric` function to the logical, character, and factor variables.**

```{r, results = "hide", warning = FALSE}
as.numeric(prob1_df$vec_logical)
as.numeric(prob1_df$vec_char)
as.numeric(prob1_df$vec_factor)
```

What happened are:  

  * Logical vector is converted to `0` (for false) and `1` (for true)  
  * Character vector is converted to `NA`  
  * Factor vector is converted to numbers
  
**\* Notice:** my factor vector is generated from a numerical vector, so here we disscuss another type of situation: what will happen if the factor vector is generated from a character vector.

```{r}
vec_factor_2 = factor(c("a", "a", "b", "b", "b", "c", "d", "e", "f", "g"))
as.numeric(vec_factor_2)
```

Here we see that characters also been converted to numbers! And the same characters designated the same number.

### step 4

**Convert character variable from character to factor to numeric:**

```{r}
as.numeric(as.factor(prob1_df$vec_char))
```

Here we can see that the character vector is first converted to factor, and since all the characters are different, they are in different factor levels.  
Then, R assigned them with different values.

<br/>

**Convert factor variable from factor to character to numeric:**

```{r}
as.numeric(as.character(prob1_df$vec_factor))
```

Here we can see that factor vector is first converted to character, and since all the factors are numerical, they are actually converted from factor (like `1`) into the same word but in type of character (like `"1"`). Then they are converted into number accordingly.

**\* Notice:** here we disscuss another type of situation: what will happen if the factor vector is generated from a character vector.

```{r}
as.numeric(as.character(vec_factor_2))
```

Here we see that the string-based factor first converted to character correspondingly, but then `NA` produced by converting characters to numbers.

## Problem 2

### step 1

**Create a data.frame.**

```{r}
prob2_df = tibble(
  x = rnorm(1000), 
  y = rnorm(1000), 
  vec_logical = ((x + y) >0), 
  vec_num = as.numeric(vec_logical), 
  vec_factor = as.factor(vec_logical)
)
```

### step 2

**Descript the vector:   **

* The dimension of the dataset is (`r dim(prob2_df)`), which means it has `r dim(prob2_df)[1]` columns and `r dim(prob2_df)[2]` rows.  
* The mean of x is `r mean(prob2_df$x)`, the median of x is `r median(prob2_df$x)`.
* The proportion of cases for which the logical vector is `TRUE`: `r length(which(prob2_df$vec_factor == 'TRUE')) / length(prob2_df$vec_factor)`.

### step 3

**Make scatterplots using `ggplot2`:**

* Make a scatterplot of `y` vs `x`, color points using the logical variable.

```{r}
ggplot(prob2_df, aes(x = x, y = y)) + 
  geom_point(aes(color = vec_logical))
```

* Make a scatterplot of `y` vs `x`, color points using the numerical variable.

```{r}
ggplot(prob2_df, aes(x = x, y = y)) + 
  geom_point(aes(color = vec_num))
```

* Make a scatterplot of `y` vs `x`, color points using the factor variable.

```{r}
ggplot(prob2_df, aes(x = x, y = y)) + 
  geom_point(aes(color = vec_factor))
```